High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-04-18 notebooks/mnist.qmd
shoot me

```{julia}
# Model parameters:
epochs = 100
batch_size = Int(round(n_obs/10))
n_hidden = 32
activation = Flux.swish
builder = MLJFlux.MLP(hidden=(n_hidden,), σ=activation)
α = [0.33,1.0,1e-1]

@@

# Joint Energy Model:
𝒟x = Uniform(-1,1)
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(𝒟x, 𝒟y, input_size=(input_dim,), batch_size=batch_size)
jem = JointEnergyClassifier(
    sampler;
    builder=builder,
    batch_size=batch_size,
    finaliser=Flux.softmax,
    loss=Flux.Losses.crossentropy,
    jem_training_params=(α=α,verbosity=10,),
    sampling_steps=20,
)

@@

```{julia}
cov = .9
conf_model = conformal_model(jem; method=:adaptive_inductive, coverage=cov)
mach = machine(conf_model, X, labels)
fit!(mach)
M = CCE.ConformalModel(mach.model, mach.fitresult)
```

```{julia}
jem = mach.model.model.jem
n_iter = 5000
_w = 1500
plts = []
neach = 10
for i in 1:10
    x = jem.sampler(jem.chain, jem.sampling_rule; niter=n_iter, n_samples=neach, y=i)
    plts_i = []
    for j in 1:size(x, 2)
        xj = x[:,j]
        xj = reshape(xj, (n_digits, n_digits))
        plts_i = [plts_i..., Plots.heatmap(rotl90(xj), axis=nothing, cb=false)]
    end
    plt = Plots.plot(plts_i..., size=(_w,0.10*_w), layout=(1,10))
    plts = [plts..., plt]
end
Plots.plot(plts..., size=(_w,_w), layout=(10,1))
```

@@

```{julia}
# Set up search:
factual_label = 9
x = reshape(counterfactual_data.X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target = 4
factual = predict_label(M, counterfactual_data, x)[1]
γ = 0.9
T = 100

@@

# Generate counterfactual using CCE generator:
generator = CCEGenerator(
    λ=[0.0,10.0], 
    temp=0.01, 
    # opt=CounterfactualExplanations.Generators.JSMADescent(η=0.5),
)
ce_conformal = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)

>>>>

```{julia}
# Model parameters:
epochs = 100
batch_size = Int(round(n_obs/10))
n_hidden = 32
activation = Flux.swish
builder = MLJFlux.@builder Flux.Chain(
    Dense(n_in, n_hidden),
    BatchNorm(n_hidden, activation),
    Dense(n_hidden, n_out),
    BatchNorm(n_out)
)
# builder = MLJFlux.MLP(
#     hidden=(
#         n_hidden,
#         n_hidden,
#         n_hidden,
#     ), 
#     σ=activation
# )
α = [1.0,1.0,1e-1]

@@

# Joint Energy Model:
𝒟x = Uniform(0,1)
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(𝒟x, 𝒟y, input_size=(input_dim,), batch_size=batch_size)
jem = JointEnergyClassifier(
    sampler;
    builder=builder,
    batch_size=batch_size,
    finaliser=x -> x,
    loss=Flux.Losses.logitcrossentropy,
    jem_training_params=(α=α,verbosity=10,),
    sampling_steps=20,
    epochs=epochs,
)

@@

```{julia}
cov = .90
conf_model = conformal_model(jem; method=:simple_inductive, coverage=cov)
mach = machine(conf_model, X, labels)
fit!(mach)
M = CCE.ConformalModel(mach.model, mach.fitresult)
```

```{julia}
jem = mach.model.model.jem
n_iter = 100
_w = 1500
plts = []
neach = 10
for i in 1:10
    x = jem.sampler(jem.chain, jem.sampling_rule; niter=n_iter, n_samples=neach, y=i)
    plts_i = []
    for j in 1:size(x, 2)
        xj = x[:,j]
        xj = reshape(xj, (n_digits, n_digits))
        plts_i = [plts_i..., Plots.heatmap(rotl90(xj), axis=nothing, cb=false)]
    end
    plt = Plots.plot(plts_i..., size=(_w,0.10*_w), layout=(1,10))
    plts = [plts..., plt]
end
plt = Plots.plot(plts..., size=(_w,_w), layout=(10,1))
display(plt)
```

@@

```{julia}
Random.seed!(1234)

# Set up search:
factual_label = 2
x = reshape(counterfactual_data.X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target = 0
factual = predict_label(M, counterfactual_data, x)[1]
γ = 0.5
T = 100

@@

# Generate counterfactual using CCE generator:
generator = CCEGenerator(
    λ=[0.0,1.0], 
    temp=0.5, 
    # opt=CounterfactualExplanations.Generators.JSMADescent(η=1.0),
)
ce_conformal = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)
